---
title: "NHS2"
output:
  html_document:
    df_print: paged
---

``` {r setup, include=FALSE}
library(tidyverse)
library(sf)
library(stringr)
library(stringi)
library(ape)
library(brms)
#install.packages("umap")
#library(umap)
```

# The Extended Natural History of Song Corpus
Summary of Analyses

``` {r load-and-filter-corpus, include=FALSE}
# load corpus
nhs2 <- read.csv('../data/metadata.csv', skip = 0, header = TRUE)

nhs2 <- nhs2 %>%
  mutate(glottocode = case_when(glottocode == ' oriy1255' ~ 'oriy1255', TRUE ~ glottocode)) %>%
  mutate(region = case_when(region == 'Northwest Coast' ~ 'Northwest Coast and California', TRUE ~ region))

n_songs <- nrow(nhs2)
```

``` {r check-missing-region-by-type, include=FALSE}
# all possible region*type categories
region_list <- unique(nhs2$region)
type_list <- unique(nhs2$type)
rt_full <- expand_grid(region_list, type_list) %>%
  rename('region'='region_list','type'='type_list')

# count all songs by region*type
reg_by_type_unfilt <- nhs2 %>%
  group_by(region, type) %>%
  dplyr::summarise(n_unfilt = n())

# fill in missing region*type rows
reg_by_type_unfilt <- left_join(rt_full, reg_by_type_unfilt) %>%
  mutate(n_unfilt = ifelse(is.na(n_unfilt), 0, n_unfilt))

reg_by_type_counts <- reg_by_type_unfilt %>%
  group_by(n_unfilt) %>%
  dplyr::summarise(number_reg_x_types = n())

```

The Extended Natural History of Song (NHS2) contains $n =$ `r toString(n_songs)` songs. All `r nrow(region_list)` HRAF regions are represented and all `r nrow(type_list)` behavioural categories are represented. Out of `r nrow(reg_by_type_unfilt)` region $\times$ type categories: `r reg_by_type_counts$number_reg_x_types[reg_by_type_counts$n_unfilt==0]` categories have 0 representative songs, `r reg_by_type_counts$number_reg_x_types[reg_by_type_counts$n_unfilt==1]` categories have only 1 representative songs, and the rest have 2 or more.


``` {r plot-geo-map, echo=FALSE, warnings=FALSE}
## TODO: add the geolocations that Thanos found + calculate centroids of family langoids as approximate location

# load map
map <- read_sf("../data/world.geo.json")

# get locations (NOTE: ~60 languages withou lat/long data)
glottolog <- read.csv('../data/languoidGlottolog.csv', skip = 0, header = TRUE) %>%
  dplyr::rename('glottocode' = 'id')

# laod up new geolocations not found in glottolog
new_geoloc <- read.csv('../data/new_geolocations.csv', skip=0, header=TRUE)

# get glottocode counts and location
lang_totals <- nhs2 %>%
  group_by(glottocode) %>%
  summarise(count = length(glottocode), region=max(region), num_reg=length(unique(region)))

# merging locations and language families
lang_wloc <- merge(lang_totals, glottolog[c('glottocode','family_id','latitude','longitude','name')])
lang_wloc_new <- merge(lang_wloc, new_geoloc[c('glottocode','latitude','longitude')], by="glottocode", all.x=TRUE)
lang_wloc_new <- lang_wloc_new %>%
  mutate(latitude.x = case_when(is.na(latitude.x) ~ latitude.y, TRUE ~ latitude.x)) %>%
  mutate(longitude.x = case_when(is.na(longitude.x) ~ longitude.y, TRUE ~ longitude.x))

# counting number of languages per family
lfam_totals <- lang_wloc %>%
  group_by(family_id) %>%
  summarise(count = sum(count)) %>%
  rename(glottocode = family_id)
  

lfam_totals <- merge(lfam_totals, glottolog[c('glottocode','name')], by="glottocode") %>% rename(family_id=glottocode)

diff_df <- setdiff(lang_totals$glottocode, lang_wloc$glottocode)

no_loc <- lang_wloc %>%
  filter(is.na(longitude))

# plot map
map_plot <- ggplot(map) + theme_void() +
  geom_sf(data=map, colour="#cccccc", fill="#dddddd") +
  geom_point(data=lang_wloc_new, aes(y=latitude.x, x=longitude.x, size=count), fill="red", shape=21, stroke=.3, colour="black") +
  #geom_point(data=lang_wloc, aes(y=latitude, x=longitude, fill=region), shape=21, stroke=.3, colour="black") +
  #facet_wrap(vars(region)) +
  scale_fill_hue(l=100, c=40)
  #labs(title="Songs in this corpus are globally distributed", caption="Locations determined by language using glottocode corpus") #+
  #theme(legend.position="none")
ggsave(map_plot, filename="map.png", height=6, width=10)
map_plot
```

Note that the locations of all but one language are found in the appropriate HRAF region. The outlier is a Ladino song (locate near Istanbull according to Glottolog) which was actually recorded in Guatemala (Maya Region). 

``` {r plot-lang-family-counts, echo=FALSE, warnings=FALSE}
lfam_totals <- lfam_totals[order(-lfam_totals$count),]

lfam_plot <- lfam_totals %>% mutate(name=fct_reorder(name,count)) %>%
  ggplot(aes(x=count, y=name, label=count)) +
  geom_col() +
  geom_text(size=3, nudge_x=12) +
  theme_bw() +
  #guides(fill="none") +
  labs(x="song count", title="# of songs by family")
ggsave(lfam_plot, filename="language_family_counts.png", height=10, width=5)
lfam_plot
```

``` {r plot-type-marginals}
type_totals <- reg_by_type_unfilt %>%
  group_by(type) %>%
  summarise(total = sum(n_unfilt))
type_totals

type_totals %>%
  ggplot(aes(x=type, y=total, fill=type, label=total)) +
  geom_col() +
  geom_label() +
  theme_bw() +
  guides(fill="none") +
  labs(y="song count", x="song type", title="Number of songs of each behavioral context")
```


``` {r plot-region-marginals}
reg_totals <- reg_by_type_unfilt %>%
  group_by(region) %>%
  summarise(reg_total = sum(n_unfilt))

region_plot <- reg_totals %>% mutate(region=fct_reorder(region,reg_total)) %>%
  ggplot(aes(y=region, x=reg_total, label=reg_total)) +
  geom_col() +
  geom_text(size=3, nudge_x=1) +
  theme_bw() +
  #guides(fill="none") +
  labs(x="song count", title="Number of songs from each region")
ggsave(region_plot, filename="region_count.png", height=6, width=10)
region_plot
```

``` {r plot-song-representation}
reg_by_type_unfilt2 <- merge(reg_by_type_unfilt, reg_totals, by='region')

reg_by_type_unfilt2 %>% mutate(region=fct_reorder(region,reg_total)) %>%
  mutate(n_unfilt = replace(n_unfilt, n_unfilt==0, NA)) %>%
  ggplot(aes(x=type, y=region, fill=!is.na(n_unfilt))) + 
  geom_bin2d() + 
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  theme_minimal() +
  labs(fill="represented", title="Which Regions and Types have Songs?")
```

``` {r plot-region-type-histograms}
reg_by_type_unfilt %>%
  ggplot(aes(x=type, y=n_unfilt, fill=type)) +
  geom_col() +
  facet_wrap(vars(region)) +
  scale_x_discrete(guide = guide_axis(n.dodge=1, angle=70)) +
  theme_bw() + theme(legend.position="bottom") +
  #guides(x="none") +
  labs(y="song count", x="", title="How many songs in each region?")
```

``` {r loading-and-comparing-acoustic-features}
audio_feats <- read.csv('../data/audioExtraction_processed.csv', skip = 0, header = TRUE)
audio_feats <- audio_feats[!audio_feats$song=="",]
length(unique(audio_feats$song))

audio_featsR <- read.csv('../data/audioExtraction_raw.csv', skip = 0, header = TRUE)
audio_featsR <- audio_featsR[!audio_featsR$song=="",]
length(unique(audio_featsR$song))

audio_featsM <- read.csv('../data/audioExtraction_raw_median.csv', skip = 0, header = TRUE)
audio_featsM <- audio_featsM[!audio_featsM$song=="",]
length(unique(audio_featsM$song))

# make tall df
af_names <- colnames(audio_feats)[-1]
af_namesM <- colnames(audio_featsM)[-1]
afP_tall <- audio_feats %>% pivot_longer(cols=all_of(af_names), names_to='feature', values_to='processed')
afR_tall <- audio_featsR %>% pivot_longer(cols=all_of(af_names), names_to='feature', values_to='raw')
afM_tall <- audio_featsM %>% pivot_longer(cols=all_of(af_namesM), names_to='feature', values_to='median')

af_tall <- merge(afP_tall,afR_tall)
afm_tall <- merge(afR_tall,afM_tall)

# include type
af_tall <- merge(af_tall, nhs2[c('song','type')],by="song")
afm_tall <- merge(afm_tall, nhs2[c('song','type')],by="song")

# COMPARING RAW vs PROCESSED
af_plot <- af_tall %>% ggplot(aes(x=raw,y=processed,color=type)) + geom_point() + 
  facet_wrap(~feature, scale="free") + theme_minimal()
#ggsave(af_plot, filename="af_compare_PvR.png", height=8, width=14)
af_plot

# COMPARING RAW MEAN vs MEDIAN
afm_plot <- afm_tall %>% ggplot(aes(x=raw,y=median,color=type)) + geom_point() + 
  facet_wrap(~feature, scale="free")# + theme_minimal()
#ggsave(afm_plot, filename="af_compare_Mean_v_Median.png", height=8, width=14)
af_plot
```

``` {r audio-features-compare}
af_tall_fam <- merge(af_tall,nhs2[c('song','glottocode','region')], by='song')
af_tall_fam <- merge(af_tall_fam,glottolog[c('glottocode','family_id')],by='glottocode')

# type proccessed
af_plot2 <- af_tall_fam %>% 
  ggplot(aes(x=processed,fill=type)) + geom_density(alpha=.3)+
  theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + facet_wrap(~feature, scale="free", ncol=2)
ggsave(af_plot2, filename="af_compare_type.png", height=18, width=8)

# type raw
af_plot3 <- af_tall_fam %>% 
  ggplot(aes(x=raw,fill=type)) + geom_density(alpha=.3)+
  theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + facet_wrap(~feature, scale="free", ncol=2)
ggsave(af_plot3, filename="af_compare_type_raw.png", height=18, width=8)

# region processed
af_plot4 <- af_tall_fam %>% 
  ggplot(aes(x=processed,fill=region)) + geom_density(alpha=.3)+
  theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(), legend.position = "top") + facet_wrap(~feature, scale="free", ncol=2)
ggsave(af_plot4, filename="af_compare_region.png", height=18, width=8)

# top 3 families proccesed
af_tall_topfam <- af_tall_fam[af_tall_fam$family_id %in% c('indo1319','aust1307','atla1278'),]
af_plot5 <- af_tall_topfam %>% 
  ggplot(aes(x=processed,fill=family_id)) + geom_density(alpha=.3) +
  theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(), legend.position = "top") +
  facet_wrap(~feature, scale="free", ncol=2)
ggsave(af_plot5, filename="af_compare_top3families.png", height=18, width=8)

# type boxplot proccessed
af_plot6 <- af_tall_fam %>% 
  ggplot(aes(y=processed,x=type, fill=type)) + geom_boxplot()+
  theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + facet_wrap(~feature, scale="free", ncol=2)
ggsave(af_plot6, filename="af_compare_type_boxplot.png", height=24, width=8)
```

``` {r load_ICC_models}
# data appended to df
get_exp_var <- function(model, feature, df) {
  # get posterior samples
  post <- posterior_samples(model)
  
  # calculate explained variance for each 
  type_sig <- post$sd_type__Intercept^2 / 
    (post$sd_glottocode__Intercept^2 + post$sd_glottocode2__Intercept^2 + post$sd_region__Intercept^2+ post$sigma^2 + post$sd_type__Intercept^2)
  tsig <- median(type_sig)
  
  phylo_sig <- post$sd_glottocode__Intercept^2 / 
    (post$sd_glottocode__Intercept^2 + post$sd_glottocode2__Intercept^2 + post$sd_region__Intercept^2+ post$sigma^2 + post$sd_type__Intercept^2)
  psig <- median(phylo_sig)

  lang_sig <- post$sd_glottocode2__Intercept^2 / 
    (post$sd_glottocode__Intercept^2 + post$sd_glottocode2__Intercept^2 + post$sd_region__Intercept^2+ post$sigma^2 + post$sd_type__Intercept^2)
  lsig <- median(lang_sig)
  
  region_sig <- post$sd_region__Intercept^2 / 
    (post$sd_glottocode__Intercept^2 + post$sd_glottocode2__Intercept^2 + post$sd_region__Intercept^2+ post$sigma^2 + post$sd_type__Intercept^2)
  rsig <- median(region_sig)
  
  resid <- post$sigma^2 / 
    (post$sd_glottocode__Intercept^2 + post$sd_glottocode2__Intercept^2 + post$sd_region__Intercept^2+ post$sigma^2 + post$sd_type__Intercept^2)
  res <- median(resid)
  
  # append values to df
  df[nrow(df)+1,1:2] <- c(feature,'type')
  df[nrow(df),3] <- tsig
  df[nrow(df)+1,1:2] <- c(feature,'phylo')
  df[nrow(df),3] <- psig
  df[nrow(df)+1,1:2] <- c(feature,'lang')
  df[nrow(df),3] <- lsig
  df[nrow(df)+1,1:2] <- c(feature,'region')
  df[nrow(df),3] <- rsig
  df[nrow(df)+1,1:2] <- c(feature,'residual')
  df[nrow(df),3] <- res
  
  return(df)
}

# create explained variance dataframe
ev_df <- data.frame(feature=character(), variable=character(),value=numeric(), stringsAsFactors=FALSE)

# load model results
# (ex_simple_lowenergy_mean)
simp_le_model <- readRDS("../data/model_fit1/simp_le_model.Rds")
ev_df <- get_exp_var(simp_le_model,'ex_simple_lowenergy_mean',ev_df)

# rename and run (ex_simple_brightness_mean)
simp_b_model <- readRDS("../data/model_fit1/simp_b_model.Rds")
ev_df <- get_exp_var(simp_b_model,'ex_simple_brightness_mean',ev_df)

# rename and run (ex_simple_roughness_mean)
simp_r_model <- readRDS("../data/model_fit1/simp_r_model.Rds")
ev_df <- get_exp_var(simp_r_model,'ex_simple_roughness_mean',ev_df)

# rename and run (ex_simple_centroid_mean)
simp_c_model <- readRDS("../data/model_fit1/simp_c_model.Rds")
ev_df <- get_exp_var(simp_c_model,'ex_simple_centroid_mean',ev_df)

# rename and run (ex_spectral_centroid_mean)
spec_cm_model <- readRDS("../data/model_fit1/spec_cm_model.Rds")
ev_df <- get_exp_var(spec_cm_model,'ex_spectral_centroid_mean',ev_df)

# rename and run (ex_spectral_centroid_std)
spec_csd_model <- readRDS("../data/model_fit1/spec_csd_model.Rds")
ev_df <- get_exp_var(spec_csd_model,'ex_spectral_centroid_std',ev_df)

# rename and run (ex_spectral_brightness_mean)
spec_bm_model <- readRDS("../data/model_fit1/spec_bm_model.Rds")
ev_df <- get_exp_var(spec_bm_model,'ex_spectral_brightness_mean',ev_df)

# rename and run (ex_spectral_brightness_std)
spec_bsd_model <- readRDS("../data/model_fit1/spec_bsd_model.Rds")
ev_df <- get_exp_var(spec_bsd_model,'ex_spectral_brightness_std',ev_df)

# rename and run (ex_spectral_spread_mean)
spec_srm_model <- readRDS("../data/model_fit1/spec_srm_model.Rds")
ev_df <- get_exp_var(spec_srm_model,'ex_spectral_spread_mean',ev_df)

# rename and run (ex_spectral_spread_std)
spec_srsd_model <- readRDS("../data/model_fit1/spec_srsd_model.Rds")
ev_df <- get_exp_var(spec_srsd_model,'ex_spectral_spread_std',ev_df)

# rename and run (ex_spectral_skewness_mean)
spec_skm_model <- readRDS("../data/model_fit1/spec_skm_model.Rds")
ev_df <- get_exp_var(spec_skm_model,'ex_spectral_skewness_mean',ev_df)

# rename and run (ex_spectral_skewness_std)
spec_sksd_model <- readRDS("../data/model_fit1/spec_sksd_model.Rds")
ev_df <- get_exp_var(spec_sksd_model,'ex_spectral_skewness_std',ev_df)

# rename and run (ex_spectral_kurtosis_mean)
spec_km_model <- readRDS("../data/model_fit1/spec_km_model.Rds")
ev_df <- get_exp_var(spec_km_model,'ex_spectral_kurtosis_mean',ev_df)

# rename and run (ex_spectral_kurtosis_std)
spec_ksd_model <- readRDS("../data/model_fit1/spec_ksd_model.Rds")
ev_df <- get_exp_var(spec_ksd_model,'ex_spectral_kurtosis_std',ev_df)

# rename and run (ex_spectral_rolloff95_mean)
spec_95m_model <- readRDS("../data/model_fit1/spec_95m_model.Rds")
ev_df <- get_exp_var(spec_95m_model,'ex_spectral_rolloff95_mean',ev_df)

# rename and run (ex_spectral_rolloff95_std)
spec_95sd_model <- readRDS("../data/model_fit1/spec_95sd_model.Rds")
ev_df <- get_exp_var(spec_95sd_model,'ex_spectral_rolloff95_std',ev_df)

# rename and run (ex_spectral_rolloff85_mean)
spec_85m_model <- readRDS("../data/model_fit1/spec_85m_model.Rds")
ev_df <- get_exp_var(spec_85m_model,'ex_spectral_rolloff85_mean',ev_df)

# rename and run (ex_spectral_rolloff85_std)
spec_85sd_model <- readRDS("../data/model_fit1/spec_85sd_model.Rds")
ev_df <- get_exp_var(spec_85sd_model,'ex_spectral_rolloff85_std',ev_df)

# rename and run (ex_spectral_spectenropy_mean)
spec_spm_model <- readRDS("../data/model_fit1/spec_spm_model.Rds")
ev_df <- get_exp_var(spec_spm_model,'ex_spectral_spectenropy_mean',ev_df)

# rename and run (ex_spectral_spectenropy_std)
spec_spsd_model <- readRDS("../data/model_fit1/spec_spsd_model.Rds")
ev_df <- get_exp_var(spec_spsd_model,'ex_spectral_spectenropy_std',ev_df)

# rename and run (ex_spectral_flatness_mean)
spec_fm_model <- readRDS("../data/model_fit1/spec_fm_model.Rds")
ev_df <- get_exp_var(spec_fm_model,'ex_spectral_flatness_mean',ev_df)

# rename and run (ex_spectral_flatness_std)
spec_fsd_model <- readRDS("../data/model_fit1/spec_fsd_model.Rds")
ev_df <- get_exp_var(spec_fsd_model,'ex_spectral_flatness_std',ev_df)

# rename and run (ex_spectral_roughness_mean)
spec_rm_model <- readRDS("../data/model_fit1/spec_rm_model.Rds")
ev_df <- get_exp_var(spec_rm_model,'ex_spectral_roughness_mean',ev_df)

# rename and run (ex_spectral_roughness_std)
spec_rsd_model <- readRDS("../data/model_fit1/spec_rsd_model.Rds")
ev_df <- get_exp_var(spec_rsd_model,'ex_spectral_roughness_std',ev_df)

# rename and run (ex_spectral_irregularity_mean)
spec_im_model <- readRDS("../data/model_fit1/spec_im_model.Rds")
ev_df <- get_exp_var(spec_im_model,'ex_spectral_irregularity_mean',ev_df)

# rename and run (ex_spectral_irregularity_std)
spec_isd_model <- readRDS("../data/model_fit1/spec_isd_model.Rds")
ev_df <- get_exp_var(spec_isd_model,'ex_spectral_irregularity_std',ev_df)

# rename and run (ex_tonal_keyclarity_mean)
tone_km_model <- readRDS("../data/model_fit1/tone_km_model.Rds")
ev_df <- get_exp_var(tone_km_model,'ex_tonal_keyclarity_mean',ev_df)

# rename and run (ex_tonal_keyclarity_std)
tone_ksd_model <- readRDS("../data/model_fit1/tone_ksd_model.Rds")
ev_df <- get_exp_var(tone_ksd_model,'ex_tonal_keyclarity_std',ev_df)

# rename and run (ex_tonal_mode_mean)
tone_mm_model <- readRDS("../data/model_fit1/tone_mm_model.Rds")
ev_df <- get_exp_var(tone_mm_model,'ex_tonal_mode_mean',ev_df)

# rename and run (ex_tonal_mode_std)
tone_msd_model <- readRDS("../data/model_fit1/tone_msd_model.Rds")
ev_df <- get_exp_var(tone_msd_model,'ex_tonal_mode_std',ev_df)

# rename and run (ex_rhythm_tempo_mean)
rhym_tm_model <- readRDS("../data/model_fit1/rhym_tm_model.Rds")
ev_df <- get_exp_var(rhym_tm_model,'ex_rhythm_tempo_mean',ev_df)

# rename and run (ex_rhythm_tempo_std)
rhym_tsd_model <- readRDS("../data/model_fit1/rhym_tsd_model.Rds")
ev_df <- get_exp_var(rhym_tsd_model,'ex_rhythm_tempo_std',ev_df)

# rename and run (ex_rhythm_attack_time_mean)
rhym_am_model <- readRDS("../data/model_fit1/rhym_am_model.Rds")
ev_df <- get_exp_var(rhym_am_model,'ex_rhythm_attack_time_mean',ev_df)

# rename and run (ex_rhythm_attack_time_std)
rhym_asd_model <- readRDS("../data/model_fit1/rhym_asd_model.Rds")
ev_df <- get_exp_var(rhym_asd_model,'ex_rhythm_attack_time_std',ev_df)

# rename and run (ex_rhythm_attack_slope_mean)
rhym_sm_model <- readRDS("../data/model_fit1/rhym_sm_model.Rds")
ev_df <- get_exp_var(rhym_sm_model,'ex_rhythm_attack_slope_mean',ev_df)

# rename and run (ex_rhythm_attack_slope_std)
rhym_ssd_model <- readRDS("../data/model_fit1/rhym_ssd_model.Rds")
ev_df <- get_exp_var(rhym_ssd_model,'ex_rhythm_attack_slope_std',ev_df)

# rename and run (ex_dynamics_rms_mean)
dyna_rm_model <- readRDS("../data/model_fit1/dyna_rm_model.Rds")
ev_df <- get_exp_var(dyna_rm_model,'ex_dynamics_rms_mean',ev_df)

# rename and run (ex_dynamics_rms_std)
dyna_rsd_model <- readRDS("../data/model_fit1/dyna_rsd_model.Rds")
ev_df <- get_exp_var(dyna_rsd_model,'ex_dynamics_rms_std',ev_df)
```

``` {r playing-w-models}
pp_check(rhym_tm_model, type="ecdf_overlay_grouped", group="type")
pp_check(rhym_am_model, type="ecdf_overlay_grouped", group="type")
```


``` {r plot-model-results}
#plot(rhym_tm_model)
post_rtm <- posterior_samples(rhym_tm_model)

ggplot(post_rtm, aes(x=sd_glottocode__Intercept, y=sd_type__Intercept)) + geom_point()
```
